---
title: "Rummikub — Function Reference"
subtitle: "R tools to understand and analyze the game of Rummikub"
format:
  html:
    toc: true
    toc-depth: 2
    theme: cosmo
    code-fold: false
    self-contained: true
    df-print: paged
---

# Overview

This project provides a set of **R functions** to analyze the board game
**Rummikub**. The goal is not to build a playable game with bots, but to
write code that helps us understand the rules, probabilities and strategies
of the game.

The functions are organized into 8 scripts, one per objective. Every script
loads the shared ones it needs with `source()`, so all of them use the same
tile representation.

## Tile representation

All functions work with `data.frame`s with two columns:

- `number`: an integer in `1:13`, or `NA` for a joker.
- `colour`: one of `"red"`, `"blue"`, `"yellow"`, `"black"`, or `"wild"`
  for a joker.

A *meld* is a `data.frame` with these same columns. A *board* is a list of
melds, or a single `data.frame` with an extra `meld_id` column.

## Repository structure

| File                              | Objective                                              |
|-----------------------------------|--------------------------------------------------------|
| `scripts/00_tile_pool.R`          | Shared tile pool definition                            |
| `scripts/01_valid_melds.R`        | 1: valid meld detection + rule of 30                   |
| `scripts/02_draw_probability.R`   | 2: probability of drawing a useful tile                |
| `scripts/03_board_verifier.R`     | 3: validate board combinations                         |
| `scripts/04_board_restructuring.R`| 4: rearrange existing combos                           |
| `scripts/05_board_visualization.R`| 5: board as `data.frame` / plot                        |
| `scripts/06_opening_analysis.R`   | 6 (analysis A): opening analysis (rule of 30, Monte Carlo)          |
| `scripts/07_runs_vs_groups.R`     | 6 (analysis B): runs vs groups (relative frequency, heatmap)        |
| `reports/function_reference.qmd`  | This document                                          |

## Setup

The following chunk loads every script (it is executed before the examples,
with the working directory inside `reports/`):

```{r setup, message = FALSE, warning = FALSE}
source("../scripts/00_tile_pool.R")
source("../scripts/01_valid_melds.R")
source("../scripts/02_draw_probability.R")
source("../scripts/03_board_verifier.R")
source("../scripts/04_board_restructuring.R")
source("../scripts/05_board_visualization.R")
source("../scripts/06_opening_analysis.R")
source("../scripts/07_runs_vs_groups.R")
```

::: {.callout-note}
`source_local()` is defined in every script. It loads a shared script from
`scripts/` by walking up from the working directory, so the scripts work no
matter where R is started from. You normally never call it directly.
:::

---

# Script 00 — Tile pool (`scripts/00_tile_pool.R`)

*Objective: a single, shared representation of the tiles used by every
other script.*

## Constant `COLOURS`

The 4 standard colours of the game.

```{r colours, message = FALSE}
COLOURS
```

## `tile_pool()`

Builds the complete pool of **106 tiles**: 104 numbered tiles (numbers 1 to
13 in 4 colours, 2 copies of each combination) plus 2 jokers.

- **Arguments:** none.
- **Value:** a `data.frame(number, colour)` with 106 rows.

```{r tile-pool, message = FALSE}
pool <- tile_pool()
nrow(pool)
head(pool, 3)
table(pool$colour)
```

## `tiles(number, colour)`

Builds a meld (a `data.frame` of tiles) from `number` and `colour` vectors.
Use `NA` / `"wild"` for jokers.

- **Arguments:**
  - `number`: integer vector (use `NA` for jokers).
  - `colour`: character vector.
- **Value:** a `data.frame(number, colour)`.

```{r tiles}
tiles(c(3, 4, 5), c("red", "red", "red"))
```

## `is_joker(number, colour)`

Tells whether a tile is a joker (`number` is `NA` or `colour == "wild"`).

- **Arguments:** `number`, `colour`: scalar tile values.
- **Value:** `TRUE`/`FALSE`.

```{r is-joker}
is_joker(NA, "wild")
is_joker(5, "red")
```

## `count_tiles(tiles)`

Builds an *inventory*: how many copies of each (number, colour) type a
collection of tiles contains.

- **Arguments:** `tiles`: a `data.frame(number, colour)`.
- **Value:** a `data.frame(number, colour, n)` with one row per present type.

```{r count-tiles}
count_tiles(tiles(c(5, 5, 5), c("red", "red", "blue")))
```

## `subtract_tiles(counter, subtract)`

Subtracts one inventory from another. Used to remove the tiles we can see
(hand + board) from the pool.

- **Arguments:**
  - `counter`: the inventory to subtract from.
  - `subtract`: the inventory to subtract.
- **Value:** the resulting inventory (rows with `n > 0` only).
- **Errors:** if a tile is not found, or if more copies are subtracted than
  exist.

```{r subtract-tiles}
after <- subtract_tiles(count_tiles(tile_pool()), count_tiles(tiles(5, "red")))
sum(after$n)   # 106 - 1 = 105 tiles left
```

---

# Script 01 — Valid melds (`scripts/01_valid_melds.R`)

*Objective 1: identify valid *runs* and *groups*, compute their points and
apply the opening rule of 30.*

## `is_run(meld)`

Is the meld a valid **run**? A run needs 3 or more consecutive tiles of the
same colour; jokers substitute for the missing numbers.

- **Arguments:** `meld`: a `data.frame(number, colour)`.
- **Value:** `TRUE`/`FALSE`.

```{r is-run}
is_run(tiles(c(3, 4, 5), c("blue", "blue", "blue")))
is_run(tiles(c(3, 4, 5), c("blue", "blue", "red")))   # different colours
is_run(tiles(c(3, 5, 7), c("blue", "blue", "blue")))   # gaps
```

## `assign_run(meld)`

Resolves the jokers of a valid run to their canonical numbers: first the
internal gaps are filled, then the run is extended to the right.

- **Arguments:** `meld`: a valid run.
- **Value:** the resolved `data.frame(number, colour)`.
- **Errors:** if `meld` is not a valid run.

```{r assign-run}
assign_run(tiles(c(5, NA, NA), c("blue", "wild", "wild")))  # -> 5,6,7
```

## `is_group(meld)`

Is the meld a valid **group**? A group needs 3 or 4 tiles with the same
number in different colours; jokers substitute for the missing colours.

- **Arguments:** `meld`: a `data.frame(number, colour)`.
- **Value:** `TRUE`/`FALSE`.

```{r is-group}
is_group(tiles(c(7, 7, 7), c("red", "blue", "black")))
is_group(tiles(c(7, 7, 7), c("red", "red", "blue")))   # repeated colour
```

## `assign_group(meld)`

Resolves the jokers of a valid group to their canonical colours (the free
colours in the standard order).

- **Arguments:** `meld`: a valid group.
- **Value:** the resolved `data.frame(number, colour)`.
- **Errors:** if `meld` is not a valid group.

```{r assign-group}
assign_group(tiles(c(7, 7, NA), c("red", "blue", "wild")))
```

## `is_valid_meld(meld)`

Is the meld a valid run or a valid group?

- **Arguments:** `meld`: a `data.frame(number, colour)`.
- **Value:** `TRUE`/`FALSE`.

```{r is-valid-meld}
is_valid_meld(tiles(c(9, 10, 11), c("red", "red", "red")))
is_valid_meld(tiles(c(2, 4, 6), c("red", "red", "red")))
```

## `meld_score(meld)`

Points of a valid meld, using the canonical resolution of jokers (a joker
counts as the tile it represents).

- **Arguments:** `meld`: a valid meld.
- **Value:** an integer with the point total.
- **Errors:** if `meld` is not valid.

```{r meld-score}
meld_score(tiles(c(9, 10, 11), c("red", "red", "red")))   # 30
meld_score(tiles(c(5, 5, 5), c("red", "blue", "yellow"))) # 15
```

## `meets_rule_30(melds)`

Applies the **opening rule**: one or more melds whose points add up to at
least 30.

- **Arguments:** `melds`: a single meld or a list of melds.
- **Value:** `TRUE`/`FALSE`.

```{r meets-rule-30}
meets_rule_30(list(
  tiles(c(8, 9, 10), c("blue", "blue", "blue")),     # 27
  tiles(c(4, 4, 4), c("red", "blue", "yellow"))      # 12
))                                                    # 27 + 12 = 39
```

---

# Script 02 — Drawing probability (`scripts/02_draw_probability.R`)

*Objective 2: probability of drawing a useful tile from the pool.*

Context: we know our hand and the tiles visible on the board; the opponent
has drawn `n_opponent` unknown tiles. Drawing is modelled as **sampling
without replacement**. For a single draw, the probability does not depend on
`n_opponent` (the drawn tile is uniform over the available pool).

## `available_pool(hand, board, n_opponent)`

Inventory of the tiles still available: the complete pool minus the hand and
the board.

- **Arguments:**
  - `hand`, `board`: `data.frame(number, colour)` or `NULL`.
  - `n_opponent`: number of unknown tiles held by the opponent.
- **Value:** an inventory `data.frame(number, colour, n)` with attribute
  `total_available` (tiles left after also removing the opponent's).
- **Errors:** if `n_opponent` is negative or larger than the pool.

```{r available-pool}
pool <- available_pool(hand = tiles(5, "red"))
sum(pool$n)                                        # 105 tiles left
attr(pool, "total_available")                      # 95 with 10 opponent tiles
available_pool(hand = tiles(5, "red"), n_opponent = 10) |>
  attr("total_available")
```

## `is_useful_tile(tile, hand)`

Default "useful" criterion. A tile is useful if it improves the hand:

- a joker is always useful;
- it matches the number of a hand tile with a different colour (group);
- it matches the colour of a hand tile and is a consecutive number (±1)
  (run).

- **Arguments:**
  - `tile`: a 1-row `data.frame(number, colour)`.
  - `hand`: the `data.frame(number, colour)` of the hand.
- **Value:** `TRUE`/`FALSE`.

```{r is-useful-tile}
is_useful_tile(tiles(6, "red"), tiles(5, "red"))   # consecutive run tile
is_useful_tile(tiles(9, "black"), tiles(5, "red")) # unrelated
```

## `useful_tiles(hand, board, n_opponent, is_useful)`

Inventory of the available tile types considered useful.

- **Arguments:** same as `available_pool()`, plus `is_useful`: the criterion
  (defaults to `is_useful_tile`).
- **Value:** an inventory `data.frame(number, colour, n)`.

```{r useful-tiles}
useful_tiles(tiles(5, "red"))
```

## `prob_draw_useful(hand, board, n_opponent, is_useful)`

Probability that the next drawn tile is useful =
useful copies / available copies.

- **Arguments:** same as `useful_tiles()`.
- **Value:** a probability in `[0, 1]`, or `NA` if there are no tiles left.

```{r prob-draw-useful}
prob_draw_useful(tiles(5, "red"))   # 12 / 105
```

---

# Script 03 — Board verifier (`scripts/03_board_verifier.R`)

*Objective 3: check that a board only contains valid combinations.*

A board is a list of melds, or a `data.frame` with a `meld_id` column.
`verify_board()` checks that (1) every meld is a valid run or group, and
(2) the tiles are physically possible (no more copies than exist in the
pool, no non-existent tiles such as a 14).

## `board_to_list(board)`

Normalizes a board to a list of melds.

- **Arguments:** `board`: a list of melds, or a `data.frame` with `meld_id`.
- **Value:** a list of `data.frame(number, colour)`.
- **Errors:** if the input format is invalid.

```{r board-to-list}
board <- list(
  tiles(c(3, 4, 5), c("red", "red", "red")),
  tiles(c(7, 7, 7), c("blue", "yellow", "black"))
)
board_to_list(board_to_dataframe(board))   # round trip
```

## `board_to_dataframe(melds)`

Converts a list of melds to a single `data.frame` with a `meld_id` column.

- **Arguments:** `melds`: a list of melds.
- **Value:** a `data.frame(meld_id, number, colour)`.

```{r board-to-dataframe}
board_to_dataframe(board)
```

## `check_physical(board_tiles)`

Checks that the given tiles could physically exist in the pool.

- **Arguments:** `board_tiles`: a `data.frame(number, colour)` with all the
  tiles on the board.
- **Value:** `list(ok, problems)`: `ok` is `TRUE`/`FALSE` and `problems` is
  a character vector of messages.

```{r check-physical}
check_physical(tiles(c(12, 13, 14), c("red", "red", "red")))
```

## `verify_board(board, details)`

Checks the whole board.

- **Arguments:**
  - `board`: a list of melds or a `data.frame` with `meld_id`.
  - `details`: if `TRUE`, return the list below instead of `TRUE`/`FALSE`.
- **Value:** `TRUE`/`FALSE`, or `list(valid, problems)` with `details = TRUE`.

```{r verify-board}
verify_board(board)
verify_board(list(tiles(c(2, 4, 6), c("red", "red", "red"))), details = TRUE)
```

---

# Script 04 — Board restructuring (`scripts/04_board_restructuring.R`)

*Objective 4: find legal ways to place tiles from the hand by adding to,
rearranging or splitting the melds already on the table.*

In Rummikub, on later turns you may break and rearrange the combos on the
table as long as everything remains valid at the end of the turn. Each move
is validated with `verify_board()`. Every move is a list with:

- `$board`: the resulting board (list of melds);
- `$hand`: the resulting hand;
- `$desc`: a readable description.

## `tile_label(tile)`

Human-readable label of a tile.

- **Arguments:** `tile`: a 1-row `data.frame(number, colour)`.
- **Value:** `"J"` for a joker, otherwise `"<number> <colour>"`.

```{r tile-label}
tile_label(tiles(NA, "wild"))
tile_label(tiles(5, "red"))
```

## `meld_type(meld)`

- **Arguments:** `meld`: a `data.frame(number, colour)`.
- **Value:** `"run"`, `"group"`, or `"invalid"`.

```{r meld-type}
meld_type(tiles(c(3, 4, 5), c("red", "red", "red")))
meld_type(tiles(c(3, 4), c("red", "red")))
```

## `resolved_meld(meld)`

The meld with its jokers resolved to their canonical values.

- **Arguments:** `meld`: a meld.
- **Value:** the resolved `data.frame`, or `NULL` if the meld is invalid.

## `move(board, hand, desc)`

Helper that builds a move list.

- **Arguments:** `board`, `hand`, `desc` (description string).
- **Value:** `list(board, hand, desc)`.

## `replace_meld(board, m, new_melds)`

Replaces meld `m` of `board` with one or more new melds.

- **Arguments:** `board`: a list of melds; `m`: the meld index;
  `new_melds`: a list of replacement melds.
- **Value:** the updated board.

## `add_to_meld(meld, tile)`

Tries to place a single tile on an existing meld: extend a run at its ends,
fill a free colour of a group, or use a joker.

- **Arguments:** `meld`: a meld; `tile`: a 1-row `data.frame`.
- **Value:** the resulting meld if valid, or `NULL` if it cannot be done.

```{r add-to-meld}
add_to_meld(tiles(c(3, 4, 5), c("red", "red", "red")), tiles(6, "red"))
add_to_meld(tiles(c(3, 4, 5), c("red", "red", "red")), tiles(9, "red"))  # gap
```

## `move_add_tile(board, hand)`

Moves that place one tile from the hand on an existing meld.

- **Arguments:** `board`, `hand`.
- **Value:** a list of moves.

```{r move-add-tile}
moves <- move_add_tile(list(tiles(c(3, 4, 5), c("red", "red", "red"))),
                       tiles(c(2, 6), c("red", "red")))
vapply(moves, function(x) x$desc, character(1))
```

## `move_swap_joker(board, hand)`

Replaces a joker on the table with a real tile from the hand and moves the
freed joker to another meld. A joker never returns to the hand.

- **Arguments:** `board`, `hand`.
- **Value:** a list of moves.

## `movable_tiles(meld)`

Indices of the tiles that can be removed from a meld leaving a valid meld.

- **Arguments:** `meld`: a meld.
- **Value:** an integer vector of row indices.

```{r movable-tiles}
movable_tiles(tiles(c(3, 4, 5, 6), c("red", "red", "red", "red")))
```

## `move_transfer_tile(board, hand)`

Moves that transfer a tile from one meld to another (both remain valid).
Does not place tiles from the hand; it prepares the ground for other moves.

- **Arguments:** `board`, `hand`.
- **Value:** a list of moves.

## `split_run(meld)`

All the ways to split a joker-free run into two valid runs of at least 3
tiles.

- **Arguments:** `meld`: a run.
- **Value:** a list of splits, each with `part1`, `part2` and `split_at`.

```{r split-run}
str(split_run(tiles(2:7, rep("red", 6))))
```

## `move_split_and_add(board, hand)`

Moves that split a run into two valid runs and add a tile from the hand to
one of the parts.

- **Arguments:** `board`, `hand`.
- **Value:** a list of moves.

```{r move-split-and-add}
moves <- move_split_and_add(list(tiles(2:7, rep("red", 6))), tiles(8, "red"))
vapply(moves, function(x) x$desc, character(1))
```

## `find_moves(board, hand)`

Collects **all** the moves above (add, swap joker, transfer, split & add)
and returns them deduplicated by description.

- **Arguments:** `board` (list of melds or `data.frame` with `meld_id`),
  `hand`.
- **Value:** a list of moves.

```{r find-moves}
moves <- find_moves(
  list(tiles(c(3, 4, 5), c("red", "red", "red"))),
  tiles(c(2, 6), c("red", "red"))
)
vapply(moves, function(x) x$desc, character(1))
```

## `move_descriptions(board, hand)`

Just the descriptions of `find_moves()` (convenient for debugging).

- **Arguments:** `board`, `hand`.
- **Value:** a character vector.

```{r move-descriptions}
move_descriptions(list(tiles(c(3, 4, 5), c("red", "red", "red"))), tiles(6, "red"))
```

---

# Script 05 — Board visualization (`scripts/05_board_visualization.R`)

*Objective 5: represent the board state as a tidy `data.frame` or a plot.*

## `order_meld(meld)`

Orders the tiles of a meld: runs by number (jokers last), groups by the
standard colour order.

- **Arguments:** `meld`: a meld.
- **Value:** the ordered `data.frame(number, colour)`.

```{r order-meld}
order_meld(tiles(c(5, 3, 4), c("red", "red", "red")))
```

## `board_to_data(board)`

Board as a tidy `data.frame` with one row per tile.

- **Arguments:** `board`: a list of melds or a `data.frame` with `meld_id`.
- **Value:** `data.frame(meld, position, number, colour, label)`, where
  `label` is the number or `"J"` for a joker.

```{r board-to-data}
board_to_data(board)
```

## `show_board(board)`

Prints a readable description of the board to the console.

- **Arguments:** `board`.
- **Value:** `invisible(NULL)`.

```{r show-board}
show_board(board)
```

## `plot_board(board, title)`

Plots the board with **ggplot2**: one cell per tile, coloured by colour and
labelled on top.

- **Arguments:** `board`; `title`: plot title.
- **Value:** a `ggplot` object (requires `ggplot2`).

```{r plot-board, warning = FALSE, fig.height = 3.5, fig.width = 7}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  plot_board(list(
    tiles(c(3, 4, 5), c("red", "red", "red")),
    tiles(c(7, 7, NA), c("blue", "yellow", "wild"))
  ))
}
```

---

# Script 06 — Opening analysis (`scripts/06_opening_analysis.R`)

*Objective 6, analysis A: the opening (rule of 30).*

Questions answered: (1) with what probability can you open using only the 14
initial tiles? (2) how many draws are needed on average to open when the
initial hand cannot reach 30?

The methods are **exact** (a branch-and-bound solver enumerates disjoint
melds) wrapped in a **Monte Carlo** simulation.

## `random_hand(k)`

- **Arguments:** `k`: number of tiles (default 14).
- **Value:** `k` random tiles from the pool, without replacement.

```{r random-hand}
random_hand(5)
```

## `prepare_hand(hand)`

Compact representation of a hand as counters per tile type.

- **Arguments:** `hand`: a `data.frame(number, colour)`.
- **Value:** `list(types, joker_idx)`:
  - `types`: `data.frame(number, colour, n, value)` with one row per present
    type (`value` is the tile number, `13` as an upper bound for jokers);
  - `joker_idx`: index of the joker row (`integer(0)` if none).

```{r prepare-hand}
prep <- prepare_hand(tiles(c(4, 5, 6, 6, 6), c("red", "red", "red", "blue", "yellow")))
prep$types
```

## `possible_melds(prep)`

Enumerates all melds (runs and groups, jokers included) formable with the
hand.

- **Arguments:** `prep`: the output of `prepare_hand()`.
- **Value:** `list(M, sums, type)`:
  - `M`: matrix (melds × types) with the tiles each meld uses;
  - `sums`: the point total of each meld;
  - `type`: `"run"` or `"group"` for each meld (a meld made of a single real
    tile plus jokers could be either, so the type is tagged at generation).

```{r possible-melds}
melds <- possible_melds(prep)
nrow(melds$M)          # number of formable melds
head(melds$sums, 8)    # their points
```

## `solve_opening(hand, target)`

Exact maximum score achievable with disjoint melds of the hand.

- **Arguments:** `hand`; `target`: with a finite value the search stops as
  soon as it is reached (ideal for the rule of 30); with `Inf` it computes
  the exact maximum.
- **Value:** an integer score (0 if no meld can be formed).

```{r solve-opening}
solve_opening(tiles(c(4, 5, 6, 4, 5, 6, NA),
                    c(rep("red", 3), rep("blue", 3), "wild")))
```

## `can_open(hand)`

Can the hand open under the rule of 30?

- **Arguments:** `hand`.
- **Value:** `TRUE`/`FALSE`.

```{r can-open}
can_open(tiles(c(9, 10, 11, NA), c("red", "red", "red", "wild")))
```

## `max_opening(hand)`

Exact maximum opening score (for the score distribution).

- **Arguments:** `hand`.
- **Value:** an integer.

```{r max-opening}
max_opening(tiles(c(3, 4, 5, 6, 7), rep("red", 5)))   # 3+4+5+6+7 = 25
```

## `simulate_draws_until_open()`

Draws tiles one at a time from the pool until the hand can open.

- **Arguments:** none.
- **Value:** the number of draws needed (0 if the 14 initial tiles already
  open).

```{r simulate-draws}
set.seed(1)
simulate_draws_until_open()
```

## Running the analysis

When launched directly (`Rscript scripts/06_opening_analysis.R`), the script
runs 5,000 random hands plus 5,000 draw simulations, prints a summary to the
console and saves the results in `results/`:

- `opening_probability.csv` — probability of opening with the 14 initial
  tiles (with 95% confidence interval).
- `opening_max_score.csv` — maximum opening score per hand.
- `opening_draws.csv` — draws needed per simulated game.
- `opening_draws_distribution.csv` — distribution of draws.
- `opening_draws_summary.csv` — summary statistics.
- `opening_score_histogram.png` — histogram of the maximum score.
- `opening_cumulative_probability.png` — cumulative probability of having
  opened after `t` draws.

---

# Script 07 — Runs vs groups (`scripts/07_runs_vs_groups.R`)

*Objective 6, analysis B: relative frequency of runs and groups in random
hands, and a heatmap of melds by tile number.*

The melds are enumerated **exactly** with `possible_melds()` from script 06,
which tags each meld as `"run"` or `"group"`. This type cannot be re-derived
from the tile matrix alone, because a meld made of a single real tile plus
jokers could be either.

## `tally_hand_melds(prep)`

Counts, for one prepared hand, the formable melds by type and by tile number.

- **Arguments:** `prep`: the output of `prepare_hand()`.
- **Value:** a list with `n_runs`, `n_groups` and the heat vectors
  `heat_run` / `heat_group` (integer(13): number of melds of each type that
  use each tile number).

```{r tally-hand}
tally_hand_melds(prepare_hand(tiles(c(4, 5, 6, 6, 6),
                                    c("red", "red", "red", "yellow", "black"))))
```

## `tally_runs_groups(n_hands, k, seed)`

Monte Carlo tally over random hands.

- **Arguments:** `n_hands`: number of hands (default 10000); `k`: tiles per
  hand (default 14); `seed`: optional seed for reproducibility.
- **Value:** a list with:
  - `per_hand`: `data.frame(n_runs, n_groups)` with one row per hand;
  - `heat`: `data.frame(number, run, group)` with the total number of melds
    of each type that use each tile number (1 to 13);
  - `n_hands`, `k`: the parameters used.

```{r tally-runs-groups}
set.seed(1)
res <- tally_runs_groups(50L)
sum(res$per_hand$n_runs)   # total runs
head(res$heat, 4)
```

## Running the analysis

When launched directly (`Rscript scripts/07_runs_vs_groups.R`), the script
runs 10,000 random hands, prints a summary to the console and saves the
results in `results/`:

- `runs_vs_groups_per_hand.csv` — runs and groups per hand.
- `runs_vs_groups_summary.csv` — aggregate statistics (relative frequency,
  mean melds per hand, dominance of each type).
- `runs_vs_groups_heatmap.csv` — melds by tile number and type
  (total and mean per hand).
- `runs_vs_groups_distribution.png` — histogram of melds per hand by type.
- `runs_vs_groups_heatmap.png` — heatmap of melds by tile number and type.

---

# Summary table

| Script | Key functions |
|--------|---------------|
| `00_tile_pool.R` | `tile_pool()`, `tiles()`, `is_joker()`, `count_tiles()`, `subtract_tiles()` |
| `01_valid_melds.R` | `is_run()`, `is_group()`, `assign_run()`, `assign_group()`, `is_valid_meld()`, `meld_score()`, `meets_rule_30()` |
| `02_draw_probability.R` | `available_pool()`, `is_useful_tile()`, `useful_tiles()`, `prob_draw_useful()` |
| `03_board_verifier.R` | `board_to_list()`, `board_to_dataframe()`, `check_physical()`, `verify_board()` |
| `04_board_restructuring.R` | `add_to_meld()`, `move_add_tile()`, `move_swap_joker()`, `move_transfer_tile()`, `move_split_and_add()`, `find_moves()`, `move_descriptions()` |
| `05_board_visualization.R` | `order_meld()`, `board_to_data()`, `show_board()`, `plot_board()` |
| `06_opening_analysis.R` | `random_hand()`, `prepare_hand()`, `possible_melds()`, `solve_opening()`, `can_open()`, `max_opening()`, `simulate_draws_until_open()` |
| `07_runs_vs_groups.R` | `tally_hand_melds()`, `tally_runs_groups()` |
| `06_opening_analysis.R` | `random_hand()`, `prepare_hand()`, `possible_melds()`, `solve_opening()`, `can_open()`, `max_opening()`, `simulate_draws_until_open()` |

Rendered from `reports/function_reference.qmd` with Quarto.
